---
title: "exploratory data analysis week 4"
author: "Radhakant Tiwari"
date: "23/05/2021"
output: word_document
---

```{r}
getwd()
setwd("D:/R FILES/exploratory data analysis week 4")

```


```{r}
getwd()
```


```{r}
NEI<- readRDS("summarySCC_PM25.rds")
SCC<- readRDS("Source_Classification_Code.rds")

tot.emissions.year <- NEI %>%  # Group by year and summarize total emissions across the board
    group_by(year) %>%
    summarize(Total.Emissions = sum(Emissions, na.rm = TRUE))

tot.emissions.year


with(tot.emissions.year, # plot data 
     plot(x = year, 
          y = Total.Emissions, 
          ylab = "Annual Emissions [Tons]", 
          xlab = "Year",
          main = "Total Annual Emissions in the US by Year",
          cex = 2,
          pch = 3,
          col = "red",
          lwd = 3))
```


```{r}
# Find delta between 2008 and 1999

tot.emissions.2008 <- tot.emissions.year[tot.emissions.year$year == 2008, 2]
tot.emissions.1999 <- tot.emissions.year[tot.emissions.year$year == 1999, 2]

delta.tot.emissions <- tot.emissions.2008 - tot.emissions.1999
delta.tot.emissions

## shows emminson is decreased
```


Have total emissions from PM2.5 decreased in the Baltimore City, Maryland (𝚏𝚒𝚙𝚜 == “𝟸𝟺𝟻𝟷𝟶”) from 1999 to 2008? Use the base plotting system to make a plot answering this question.

```{r}
tot.emissions.baltimore <- NEI %>%
        subset(fips == "24510") %>%
        group_by(year) %>%
        summarize(Total.Emissions.Baltimore = sum(Emissions, 
                                                  na.rm = TRUE))
    
with(tot.emissions.baltimore, 
     plot(x = year, 
          y = Total.Emissions.Baltimore, 
          ylab = "Total Annual Emissions [Tons]", 
          xlab = "Year",
          main = "Total Annual Emissions in Baltimore by Year",
          cex = 2,
          pch = 3,
          col = "red",
          lwd = 3))
```

Of the four types of sources indicated by the 𝚝𝚢𝚙𝚎 (point, nonpoint, onroad, nonroad) variable, which of these four sources have seen decreases in emissions from 1999–2008 for Baltimore City?

Which have seen increases in emissions from 1999–2008? Use the ggplot2 plotting system to make a plot answer this question.

```{r}
tot.emissions.type <- NEI %>% 
    subset(fips == "24510") %>%
    group_by(year, type) %>%
    summarize(Total.Emissions.Type = sum(Emissions, na.rm = TRUE))

emissions.type <- ggplot(data = tot.emissions.type, aes(year, Total.Emissions.Type))

emissions.type <- emissions.type + 
    geom_point(color = "red", 
               size = 4, 
               alpha = 1/3) + 
    facet_grid(. ~ type) +
    xlab("Year") +
    ylab("Total Emissions [Tons]") +
    ggtitle("Total Annual Emissions in Baltimore by Year")

emissions.type

```

Across the United States, how have emissions from coal combustion-related sources changed from 1999–2008?

To find the contributions to emissions from coal combustion, we should merge the NEI and SCC datasets to see the EI.Sector column, which will be filtered for words containing “coal”. The Short.Name column also contains this information; however, it includes contributions from coal that are not from combustion.
```{r}
SCC.coal.comb <- SCC[grep("[Cc]oal",SCC$EI.Sector),]
NEI.sub.coal <- subset(NEI, 
                       NEI$SCC %in% SCC.coal.comb$SCC)


NEI.coal.comb <- merge(x = NEI.sub.coal, 
                       y = SCC, 
                       by.x = "SCC", 
                       by.y = "SCC")

NEI.coal.comb.tot <- NEI.coal.comb %>% 
    group_by(year) %>%
    summarize(Total.Coal.Comb = sum(Emissions, na.rm = TRUE))

NEI.coal.comb.plot <- ggplot(NEI.coal.comb.tot, aes(year, Total.Coal.Comb))

NEI.coal.comb.plot <- NEI.coal.comb.plot + 
    geom_point(color = "red", 
               size = 4, 
               alpha = 1/3) + 
    xlab("Year") +
    ylab("Total Emissions [Tons]") +
    ggtitle("Total Annual Coal Combustion Emissions in the US")

NEI.coal.comb.plot
```


```{r}
NEI.coal.comb.tot.2008 <- NEI.coal.comb.tot[NEI.coal.comb.tot$year == 2008, 2]
NEI.coal.comb.tot.1999 <- NEI.coal.comb.tot[NEI.coal.comb.tot$year == 1999, 2]

NEI.coal.comb.delta <- NEI.coal.comb.tot.2008 - NEI.coal.comb.tot.1999
NEI.coal.comb.delta
```

How have emissions from motor vehicle sources changed from 1999–2008 in Baltimore City?

```{r}
vehicle.scc <- SCC[grep("[Vv]eh", SCC$Short.Name), ]

emissions.motor.baltimore <- NEI %>% 
    subset(fips == "24510" & NEI$SCC %in% vehicle.scc$SCC) %>%
    merge(y = vehicle.scc, by.x = "SCC", by.y = "SCC") %>%
    group_by(year) %>%
    summarize(Vehicle.Emissions.Type = sum(Emissions, na.rm = TRUE))

emissions.motor.baltimore.plot <- ggplot(emissions.motor.baltimore, aes(year, Vehicle.Emissions.Type)) +
    geom_point(color = "red", 
               size = 4, 
               alpha = 1/3) + 
    xlab("Year") +
    ylab("Total Emissions [Tons]") +
    ggtitle("Total Annual Vehicle Emissions in Baltimore City")

emissions.motor.baltimore.plot
```

```{r}
emissions.motor.baltimore.2008 <- emissions.motor.baltimore[emissions.motor.baltimore$year  == 2008, 2]
emissions.motor.baltimore.1999 <- emissions.motor.baltimore[emissions.motor.baltimore$year  == 1999, 2]

delta.baltimore <- emissions.motor.baltimore.2008 - emissions.motor.baltimore.1999
delta.baltimore
```
Compare emissions from motor vehicle sources in Baltimore City with emissions from motor vehicle sources in Los Angeles County, California (𝚏𝚒𝚙𝚜 == “𝟶𝟼𝟶𝟹𝟽”). Which city has seen greater changes over time in motor vehicle emissions?

```{r}

vehicle.scc <- SCC[grep("[Vv]eh", SCC$Short.Name), ]

emissions.motor.la <- NEI %>% 
    subset(fips == "06037" & NEI$SCC %in% vehicle.scc$SCC) %>%
    merge(y = vehicle.scc, by.x = "SCC", by.y = "SCC") %>%
    group_by(year) %>%
    summarize(Vehicle.Emissions.Type = sum(Emissions, na.rm = TRUE))

emissions.motor.baltimore2 <- cbind(emissions.motor.baltimore, "City" = rep("Baltimore", 4))
emissions.motor.la2 <- cbind(emissions.motor.la, "City" = rep("LA", 4))

emissions.motor.comp <- rbind(emissions.motor.baltimore2, emissions.motor.la2)

emissions.motor.comp.plot <- ggplot(emissions.motor.comp, aes(year, Vehicle.Emissions.Type, col = City)) +
    geom_point(size = 4, 
               alpha = 1/3) +
    xlab("Year") +
    ylab("Total Emissions [Tons]") +
    ggtitle("Comparison of Total Annual Vehicle Emissions in Baltimore and Los Angeles")

emissions.motor.comp.plot

```


```{r}
emissions.motor.la.2008 <- emissions.motor.la[emissions.motor.la$year  == 2008, 2]
emissions.motor.la.1999 <- emissions.motor.la[emissions.motor.la$year  == 1999, 2]

 
delta.la <- emissions.motor.la.2008 - emissions.motor.la.1999
abs(delta.la) > abs(delta.baltimore)

delta.la
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


